Visual Interrogation of Attention-Based Models for Natural Language Inference and Machine Comprehension

### 2018-05-31 EMNLP-demo-paper/conclusion.tex
update result

Even though we design the individual components with versatility in mind, due to a large number of variants of attention networks, it is hard to ensure the current interface is compatible with all the available configurations.
%
In the future, we plan to improve upon the current attention interface and expand the visualization component further to handle tasks such neural machine translation and more.

>>>>

Even though we design the individual components with versatility in mind, due to a large number of variants of attention networks, it is hard to ensure the current interface is compatible with all the available configurations.
%
In the future, we plan to improve upon the current attention interface, release the library as an open source package, and expand the visualization components to handle tasks such neural machine translation and more.

### 2018-05-31 EMNLP-demo-paper/result.tex
update result

%\begin{figure*}[t]
%\centering
%\vspace{-2mm}
% \includegraphics[width=1.0\linewidth]{NLI_MC_interface}
%  \vspace{-6mm}
% \caption{
%Illustration of different configurations for the natural language inference and machine comprehension tasks.
%}
%\label{fig:pipelineUpdate}
%\end{figure*}
\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{NLIexample}
  \vspace{-6mm}
 \caption{
An illustration of the attention editing process.
The dependency structure is shown in (a), where the two ``green'' are decorating different nouns.
By removing the ``wrong'' alignment in (b), the originally prediction \emph{entailment} is corrected to \emph{netural} in (c).
}
  \vspace{-3mm}
\label{fig:NLIexample}
\end{figure*}

@@

As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pairs (\textbf{P}:``A woman in a green jacket is drinking tea.'' \textbf{H}:``A woman is drinking green tea.'') is predicted to be entailment, which is incorrect.
By examining the attention, we can see the word \emph{green} in ``\emph{green} jacket" is aligned to the \emph{green} in ``\emph{green} tea". However, these two ``green'' decorate different nouns, which lead to the wrong prediction. The grammar structure is captured by the dependency tree, i.e., \emph{green} in \textbf{P} is attached to ``jacket''. However, the model does not has access to the grammar information and mistakenly assume the two \emph{greens} are describing the same thing (thus predict \emph{entailment}). 
%
To correct the mistake, we can edit the attention value and remove the align between these two "green" (see Figure~\ref{fig:NLIexample}(b)).
As expected, the prediction label is corrected (\emph{neutral}) (see in Figure~\ref{fig:NLIexample}(c)).

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{MC_interface}
  \vspace{-6mm}
 \caption{
In the machine comprehension visualization interface (a), the $p1$, $p2$ color bar illustrate the predicted start and end index for the answer (deeper the red the higher the probability).
The answer to the question is found in (a1), where the ``Two'' is aligned with the ``many'' in the question. However, the number 9 is also aligned with many, which may potentially lead to confusions.
%
We can evaluate the robustness of the prediction by perturbing the question sentence (b1, b3). As illustrated in (b1, b2) by removing the word ``partial'' the model still find the correct answer (albeit different, as the sentence perturbation changes the exact meaning of the question). 
}
\label{fig:MCexample}
\end{figure*}

@@

As illustrated in Figure~\ref{fig:MCexample}, we show \emph{att2} in the overview attention colored bars (yellow-green-blue colormap), in which each rectangular corresponds to one sentence. We can then focus on individual sentences by click on the rectangular (see Figure~\ref{fig:MCexample}).
%The proposed visualization help reveal the potential alignment issues in the machine comprehension model.
The $p1$, $p2$ colored bars (white-red colormap) illustrate the predicted start and end index for the answer (deeper the red the higher the probability). The span of text defined by the start and end index with the highest probability is the answer.
%
As we can see, there are two spans of text that correspond to higher probabilities. We can examine the corresponding sentence in (a3) and (b2). Both sentences exhibit excellent alignment with the question. The spans with high probability are ``Two'' and ``No'' (highlight in orange), which both align well with many.
Another interesting observation is that the number ``9'' (in ``Falcon 9') is also aligned with ``many'', yet in the prediction both ``Two'' and ``No'' have significantly higher probability.
One possible explanation is the word ``Two'' has much higher \emph{att2} attention value compare to ``9'', which may contribute to the current outcome.

>>>>

%\begin{figure*}[t]
%\centering
%\vspace{-2mm}
% \includegraphics[width=1.0\linewidth]{NLI_MC_interface}
%  \vspace{-6mm}
% \caption{
%Illustration of different configurations for the natural language inference and machine comprehension tasks.
%}
%\label{fig:pipelineUpdate}
%\end{figure*}
\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{NLIexample}
  \vspace{-6mm}
 \caption{
An illustration of the attention editing process.
The dependency structure is shown in (a), where the two ``green'' are decorating different nouns.
By removing the ``wrong'' alignment in (b), the originally prediction \emph{entailment} is corrected to \emph{netural} in (c).
}
\vspace{-2mm}
\label{fig:NLIexample}
\end{figure*}

@@

As illustrated in Fig.~\ref{fig:NLIexample}(a), the input sentence pairs (\textbf{P}:``A woman in a green jacket is drinking tea.'' \textbf{H}:``A woman is drinking green tea.'') is predicted to be entailment, which is incorrect.
By examining the attention, we can see the word \emph{green} in ``\emph{green} jacket" is aligned to the \emph{green} in ``\emph{green} tea". However, these two ``green'' decorate different nouns, which lead to the wrong prediction. The grammar structure is captured by the dependency tree, i.e., \emph{green} in \textbf{P} is attached to ``jacket''. However, the model does not has access to the grammar information and mistakenly assume the two \emph{greens} are describing the same thing (thus predict \emph{entailment}). 
%
To correct the mistake, we can edit the attention value and remove the align between these two "green" (see (c)(b)).
As expected, the prediction label is corrected (\emph{neutral}).

@@

\begin{figure*}[t]
\centering
\vspace{-2mm}
 \includegraphics[width=1.0\linewidth]{MC_interface}
  \vspace{-6mm}
 \caption{
In the machine comprehension visualization interface (a), the $p1$, $p2$ colored bar (in $a3$) illustrate the predicted start and end index of the answer in the context (the deeper the red the higher the probability). The most likely answer is shown in ($a1$). The global and local attention are visualized by ($a2$, $a3$).
%
We can evaluate the robustness of the prediction by perturbing the question sentence ($b1$, $b3$). As illustrated in ($b1$, $b2$) by removing the word ``partial'' the model still find the correct answer (albeit different, as the sentence perturbation changes the exact meaning of the question). 
}
\label{fig:MCexample}
\end{figure*}

As illustrated in Figure~\ref{fig:MCexample}, we show \emph{att2} as the colored bars with \emph{yellow-green-blue} colormap. Each rectangular bar corresponds to one sentence. The user can focus on individual sentences by click on the bar (see Figure~\ref{fig:MCexample}).
%The proposed visualization help reveal the potential alignment issues in the machine comprehension model.
The $p1$, $p2$ colored bars (\emph{white-red} colormap) illustrate the predicted start and end index for the answer (deeper the red the higher the probability). The span of text defined by the start and end index with the highest probability is the answer.
%
As we can see in Figure~\ref{fig:MCexample} ($a3$), there are two spans of text that correspond to higher probabilities. We can examine the corresponding sentence in (a3) and (b2). Both sentences exhibit excellent alignment with the question. The spans with high probability are ``Two'' and ``No'' (highlight in orange), which both align well with the word ``many''.
%Another interesting observation is that the number ``9'' (in ``Falcon 9') is also aligned with ``many'', yet in the prediction both ``Two'' and ``No'' have significantly higher probability.
We can explore the robustness of the model by examining how the prediction varies when the question is perturbed.
%
As illustrated in ($b1$, and $b2$), the perturbed sentence does not include the ``word'' partial, which lead to a different yet still correct answer ``No'', which also exhibits high probability for the original question.
%
The user can manually edit the text directly. As shown in ($b3$), the model still produces the correct answer when changing the question from ``how many'' to ``which''.
%One possible explanation is the word ``Two'' has much higher \emph{att2} attention value compare to ``9'', which may contribute to the current outcome.
